---
title: "Spotify Project Markdown"
author: "Andrew Gregory"
date: '2022-06-29'
output: pdf_document
---

```{r}
library("tidyverse")
library("ggExtra")
library("ggthemes")
library("ggpubr")
library("magrittr")
library("ggcorrplot")
library(ggExtra)
```



Data loading
```{r}
# Load data for the 2010's
df10 <- read_csv(file = "Project_Data/dataset-of-10s.csv")
```
Feature Engineering
```{r}
# Change to seconds for the sake of interpretability
df10 %<>%
  mutate(duration_s = duration_ms / 1000)
# Change these columns to factors
df10 %<>%
  mutate(across(c(mode, key, time_signature), ~ factor(.x)))
# List factor column
df10 %>%
  select(where(is.factor))
# Delete uninformative columns
df10 %<>%
  select(-track, -artist, -uri, -target, -duration_ms)
```
Define Spotify green for the sake of plotting
```{r}
spotify_green <- "#1DB954"
```


Numeric plots
```{r}
numeric_labels <- c(
  `sections` = "Sections in Song", `chorus_hit` = "Time Until Chorus Hit",
  `duration_s` = "Song Duration", `tempo` = "Tempo",
  `valence` = "Valence", `liveness` = "Liveness", `instrumentalness` = "Instrumentalness",
  `acousticness` = "Acousticness", `speechiness` = "Speechiness", `loudness` = "Loudness",
  `energy` = "Energy", `danceability` = "Danceability"
)
df10 %>%
  select(where(is.numeric)) %>%
  gather() %>%
  ggplot(aes(value)) +
  geom_histogram(fill = spotify_green) +
  ylab("Frequency") +
  xlab("Feature Value") +
  theme_tufte() +
  facet_wrap(~key, scales = "free", labeller = as_labeller(numeric_labels))
```


Discrete plots
```{r}
factorLabels <- c(`mode` = "Mode", `key` = "Key", `time_signature` = "Time Signature")
# To do: Figure out why time_signature is considered a different data type
# df10 %>%
#   select(mode, key, time_signature) %>%
#   gather() %>%
#   ggplot(aes(value)) +
#   geom_bar(fill = SpotifyGreen) +
#   facet_wrap(~key, scales = 'free')
# par(mfrow = c(3,1))
modeP <- df10 %>%
  ggplot(aes(x = mode)) +
  geom_bar(fill = spotify_green) +
  theme_tufte() +
  xlab("Mode") +
  ylab("Count")
keyP <- df10 %>%
  count(key) %>%
  ggplot(aes(x = fct_reorder(key, n), y = n)) +
  geom_col(fill = spotify_green) +
  theme_tufte() +
  ylab("Count") +
  xlab("Key")
tsP <- df10 %>%
  count(time_signature) %>%
  ggplot(aes(x = fct_reorder(time_signature, n), y = n)) +
  geom_col(fill = spotify_green) +
  theme_tufte() +
  ylab("Count") +
  xlab("Time Signature")
ggarrange(modeP, tsP, keyP,
  ncol = 2, nrow = 2, align = "hv",
  widths = c(2, 1), heights = c(1, 2),
  common.legend = TRUE
)
```



Relationships between two variables
```{r}
df10 %>%
  select(where(is.numeric)) %>%
  cor(method = "spearman") %>%
  ggcorrplot(colors = c("Black", "Grey", SpotifyGreen)) +
  scale_x_discrete(labels = c(
    "Sections in Song", "Time Until Chorus Hit", "Duration of Song in Seconds", "Tempo",
    "Valence", "Liveness", "Instrumentalness", "Acousticness", "Speechiness", "Loudness",
    "Energy", "Danceability"
  )) +
  scale_y_discrete(labels = c(
    "Sections in Song", "Time Until Chorus Hit", "Duration of Song in Seconds", "Tempo",
    "Valence", "Liveness", "Instrumentalness", "Acousticness", "Speechiness", "Loudness",
    "Energy", "Danceability"
  ))
```

Looks like danceability and Energy have some interesting correlation

```{r}
(df10 %>%
  ggplot(aes(danceability, energy)) +
  geom_point(color = spotify_green) +
  geom_smooth(method = "lm", color = "black") +
  theme_tufte()) %>%
  ggMarginal(type = "hist")
```

Chorus hit and Valence it looks like
```{r}

(df10 %>%
  ggplot(aes(log(chorus_hit + 1), valence)) +
  geom_point(color = SpotifyGreen, size = .2) +
  geom_smooth(color = "black", method = "lm") +
  xlab("Log Time Until Chorus Hit in Seconds") +
  ylab("Valence") +
  theme_tufte()) %>%
  ggMarginal(type = "hist")
```
Looks like that high correlation was a result of some high leverage outliers.

Time until chorus and duration of song
```{r}
(df10 %>%
  ggplot(aes(log(chorus_hit + 1), log(duration_s))) +
  geom_point(color = SpotifyGreen, size = .2) +
  geom_smooth(color = "black") +
  xlab("Log Time Until Chorus Hit in Seconds") +
  ylab("Log Duration of Time in Seconds") +
  theme_tufte()) %>%
  ggMarginal(type = "histogram")
```

Liveness and Acoustiness
```{r}
(df10 %>%
  ggplot(aes(log(liveness + 1), log(acousticness + 1))) +
  geom_point(color = spotify_green) +
  geom_smooth(color = "black") +
  xlab("Liveness") +
  ylab("Acousticness") +
  theme_tufte()
) %>%
  ggMarginal(type = "hist")
```










